\chapter{Acknowledgements}\label{acknowledgements}

\begin{multicols}{2}
At \href{http://www.rcc.mass.edu}{RCC}, introductory (general) biology
is split into two courses (SCI103, Biology I, and SCI104, Biology II)
which are taught over two semesters. The two courses were developed in
the 1970s by Prof.~Georgia Whitman essentially representing botany
(Biology I) and zoology (Biology II). Some 20 years ago, Prof.~Kyrsis
Rodriguez who served as the biology course coordinator until her
retirement reduced the botany content and introduced molecular aspects
of biology.

In 2015, the Massachusetts Department of Higher Education with its
\href{http://www.mass.edu/masstransfer/}{MassTransfer Pathways}
initiative prompted a statewide critical evaluation of the structure and
content of foundational courses, such as introductory biology, in order
to make these courses eligible for transfer across the Commonwealth
amongst public higher education institutions. In 2016, we have been
fortunate to receive a large grant from the
\href{http://www.masslifesciences.com}{Massachusetts Life Sciences
Center} to upgrade and modernize our science labs. We have taken this
opportunity to re-structure our courses along the lines of current
practice of introductory biology teaching such that the Biology I course
introduces the molecular and cellular basis of life, while the Biology
II course is focused on the basics of organismal biology. Overall the
proposed changes will expose our students to modernized (up-to-date)
concepts of biology and thus make them more competitive.

This laboratory manual was inspired by
``\href{https://www.morton-pub.com/catalog/biology/encounters-life-7e}{Encounters
with Life}'' by Larry J. Scott and Hans F. E. Wachtmeister published by
the Morton Publishing Company in Englewood CO. We used that manual at
RCC for many years. Progress in the biological sciences is, however,
relentless and so over the years, our requirements changed, and we felt
that it was best to put together a new manual specifically tailored to
our needs.

The creation of this manual was greatly facilitated and owes a major
debt to \href{https://www.wikipedia.org}{Wikipedia} and its large number of voluntary contributors. I very
liberally copied from many Wikipedia pages and then remixed, edited,
adapted and added text. With your continued support and help this manual
can only get better over time. I urge you to email me with your
criticisms and suggestions at
\href{mailto:nsucher@rcc.mass.edu}{\nolinkurl{nsucher@rcc.mass.edu}}.
This manual is made available as an open educational resource under
\href{https://creativecommons.org/licenses/by-sa/3.0/deed.en}{Creative
Commons Attribution-Share Alike 3.0 Unported} United States License for
others to do as I did and improve and adapt to specific requirements.

I wish to thank Dr.~Hillel Sims, Dean of STEM, for the laboratory safety
and microscope videos. Last but not least, I want to thank all lab
technicians, my teaching colleagues and our students who work together
to translate mere words into an exciting laboratory experience for all
of us at RCC.

\end{multicols}
